### 2019-02-03 Seneca.tex
mz - spell check

For parallel function invocation, Seneca integrates 
Celery~\cite{ref:celery}. Celery is an asynchronous task queue 
that uses distributed message passing. Celery workers are processes 
that take tasks from the queue, execute the tasks with the auguments specified, 
and store the result that is returned 
in a database (we use Redis~\cite{ref:redis} in our prototype).

>>>>

For parallel function invocation, Seneca integrates 
Celery~\cite{ref:celery}. Celery is an asynchronous task queue 
that uses distributed message passing. Celery workers are processes 
that take tasks from the queue, execute the tasks with the arguments specified, 
and store the result that is returned 
in a database (we use Redis~\cite{ref:redis} in our prototype).

### 2019-02-03 abstract.tex
mz - update abstract
mz - spell check

The goal of our work is to simplify and expedite the 
construction and evaluation of machine learning models 
using autoscaled cloud computing systems and services.
To enable this, we develop an open source system, called Seneca, that leverages
the serverless programming model and its implementation in Amazon Web Services
(AWS) Lambda.  
Seneca takes a machine learning application, dataset, and a list of 
possible hyperparameter options.  It automatically constructs an AWS Lambda
handler that invokes the application, ingresses and splits the dataset into training
and testing portions, and constructs, tests, and evaluates (i.e. scores) a machine
learning model for a given set of hyperparameter values.  Seneca orchestrates
this execution in parallel for all possible combinations of the hyperparameters specified.
It then returns the configuration (and model) that results in the best score 
to the user.

We present the design and implementation of Seneca and introduce an extension to the
system that automaticallly optimizes memory use by the functions.  
Our empirical evaluation using
multiple machine learning applications for regression and classification (using
the models for prediction and not explanatory value), shows that Seneca is able to quickly
identify the best performing model,
for the datasets and algorithms that we study. 
Moreover, our memory optimization reduces the cost of using
Seneca by 10-30\% for the applications studied.

>>>>

The goal of our work is to simplify and expedite the 
construction and evaluation of machine learning models 
using autoscaled cloud computing systems and services.
To enable this, we develop an open source system, called Seneca, that leverages
the serverless programming model and its implementation in Amazon Web Services
(AWS) Lambda.  
Seneca takes a machine learning application, dataset, and a list of 
possible hyperparameter options.  It automatically constructs an AWS Lambda
handler that ingresses and splits the dataset into training
and testing portions, and constructs, tests, and evaluates (i.e. scores) a machine
learning model for a given set of hyperparameter values.  Seneca orchestrates
this execution in parallel for all possible combinations of the hyperparameters specified.
It then returns the configuration (or model) that results in the best score 
to the user.

We present the design and implementation of Seneca and introduce an extension to the
system that automatically optimizes memory use by the functions.  
Our empirical evaluation using
multiple machine learning applications for regression and classification (using
the models for prediction and not explanatory value), shows that Seneca is able to quickly
identify the best performing model,
for the datasets and algorithms that we study. 
Moreover, our memory optimization reduces the cost of using
Seneca by 10-30\% for the applications studied.

### 2019-02-03 conclusion.tex
mz - small tweaks and typos
A number of minor fixes to improve clarity and some spelling fixes.
mz - spell check

We present the design, implementation, and cost optimization for Seneca.
The optimization automaticallly optimizes function memory use to reduce the cost
of AWS Lambda use.  Our empirical evaluation using
multiple applications for regression and classification (using
the models for prediction and not explanitory value), shows that Seneca is able to quickly
identify the best performing model for the applications and datasets that we consider.  
We also show that Seneca memory optimization reduces the cost of using
Seneca by 10-30\% for the applications studied.

>>>>

We present the design, implementation, and cost optimization for Seneca.
The optimizer automatically optimizes function memory use to reduce the cost
of AWS Lambda use.  Our empirical evaluation using
multiple applications for regression and classification (using
the models for prediction and not explanatory value), shows that Seneca is able to quickly
identify the best performing model for the applications and datasets that we consider.  
We also show that Seneca memory optimization reduces the cost of using
Seneca by 10-30\% for the applications studied.

### 2019-02-03 intro.tex
mz - spell check

To date however, cloud use has been limited to large-scale 
implementations of enterprise services, such as
object storage, databases, queuing systems, and web and application servers.
The reason for this is that it is very
challenging, even for expert users, to configure complex
distributed systems for application use, 
and to leverage the elasticity (auto-scaling) that
clouds offer. To address this challenge, cloud providers 
have started to offer cloud programming and execution environments
under the moniker of \textit{serverless} 
computing~\cite{ref:jonas2017occupy, ref:onesteptwostep, ref:peeking}.  Serverless
systems automatically configure, manage, and scale applications
to significantly simplify and broaden participation in (by less expert
users) the use of cloud systems.
%xxx: http://shivaram.org/publications/pywren-socc17.pdf
%yyy: http://cidrdb.org/cidr2019/papers/p119-hellerstein-cidr19.pdf
%zzz: L. Wang, M. Li, Y. Zhang, T. Ristenpart, and M. Swift. “Peek- ing Behind the Curtains of Serverless Platforms”. In: USENIX ATC. Boston, MA: USENIX Association, 2018

Using the serverless model, applications developers upload 
arbitrary computations in high level languages
as stateless functions to these cloud-hosted services, along with a
specification conditions under which each
function should be triggered.  Functions are invoked (triggered)
automatically by the cloud in response to updates to other cloud services
(e.g. storage, queues, notification services, among others). They
communicate, persist, and access data only 
via their inputs or shared, remote storage.  
As a result, serverless applications are inherently elastic and can 
implement highly concurrent and parallel tasks.
In public clouds, users pay a small fee for the resources their 
functions use during execution, resulting in very low cost cloud use.
Although now available from all public cloud providers and as open 
source for private cloud systems, 
AWS Lambda~\cite{ref:awslambdadg} was the first and is now the 
most popular and most advanced serverless public cloud service 
available today.
%awslambda: https://docs.aws.amazon.com/lambda/latest/dg/lambda-dg.pdf

In this work, we investigate the efficacy of using AWS Lambda
for tuning machine learning applications.
To date, Lambda has not be heavily used for training and evaluating
machine learning models because of a concern that 
doing so will result in high overhead (i.e. be costly) because
of the stateless nature of serverless functions~\cite{ref:onesteptwostep}.
At the same time, identifying the ``best'' configuration for advanced
machine learning models is challenging given the large number of configuration
options, i.e., hyperparameters, typical of models today.
Hyperparameters govern the learning process of machine learning applications.
Given that parameter sweeps are embarrassingly parallel, we believe
that such tuning is a good fit for the serverless model.
To investigate this potential, its overhead, and to simplify the 
use of Lambda for training, testing, and evaluation machine learning models, 
we design and develop a new system and toolset called \textit{Seneca}.

Specifically, Seneca implements, packages, and deploys 
machine learning applications as stateless functions to AWS Lambda.
It then orchestrates exhaustive evaluation of hyperparameter settings
to identify the best performing model (for a given dataset) by
comparing the error and accuracy across models.  We consider
prediction accuracy (as opposed to explanatory power) as the
scoring metric (mean squared error for regression 
and accuracy percent for classification), to avoid overfitting.
Users present Seneca with their application, a range of values for 
each hyperparameter (or the default can be used), and a representative dataset.
Seneca produces, tests, and evaluates models for all combinations 
of hyperparameters and returns to the user
the set of parameters (or the model itself) that produces
the best cross-validation score (e.g. lowest squared error or highest
accuracy) for the dataset. Users can employ this model for other datasets
(with Seneca if desired) without retraining the model to amortize the cost
of Seneca further.

>>>>

To date however, cloud use has been limited to large-scale 
implementations of enterprise services, such as
object storage, databases, queuing systems, and web application servers.
The reason is that it is very
challenging, even for expert users, to configure complex
distributed systems for application use, 
and to leverage the elasticity (auto-scaling) that
clouds offer. To address this challenge, cloud providers 
have started to offer cloud programming and execution environments
under the moniker of \textit{serverless} 
computing~\cite{ref:jonas2017occupy, ref:onesteptwostep, ref:peeking}.  Serverless
systems automatically configure, manage, and scale applications
to significantly simplify and broaden participation (by less expert
users) in the use of cloud systems.
%xxx: http://shivaram.org/publications/pywren-socc17.pdf
%yyy: http://cidrdb.org/cidr2019/papers/p119-hellerstein-cidr19.pdf
%zzz: L. Wang, M. Li, Y. Zhang, T. Ristenpart, and M. Swift. “Peek- ing Behind the Curtains of Serverless Platforms”. In: USENIX ATC. Boston, MA: USENIX Association, 2018

Using the serverless model, applications developers upload 
arbitrary computations in high level languages
as stateless functions to these cloud-hosted services, along with a
specification conditions under which each
function should be triggered.  Functions are invoked (triggered)
automatically by the cloud in response to external events or updates to other cloud services
(e.g. storage, queues, notification services, among others). They
communicate, persist, and access data only 
via their inputs or shared, remote storage.  
As a result, serverless applications are inherently elastic and can 
implement highly concurrent and parallel tasks.
In public clouds, users pay a small fee for the resources their 
functions use during execution, resulting in very low cost cloud use.
Although now available from all public cloud providers and as open 
source for private cloud systems, 
AWS Lambda~\cite{ref:awslambdadg} was the first and is now the 
most popular and most advanced serverless public cloud service 
available today.
%awslambda: https://docs.aws.amazon.com/lambda/latest/dg/lambda-dg.pdf

In this work, we investigate the efficacy of using AWS Lambda
for tuning machine learning applications.
To date, Lambda has not be heavily used for training and evaluating
machine learning models because of a concern that 
doing so will result in high overhead (i.e. be costly) because
of the stateless nature of serverless functions~\cite{ref:onesteptwostep}.
At the same time, identifying the ``best'' configuration for advanced
machine learning models is challenging given the large number of configuration
options (i.e. hyperparameters) typical for models today.
Hyperparameters govern the learning process of machine learning applications.
Given that parameter sweeps are embarrassingly parallel, we believe
that such tuning is a good fit for the serverless model.
To investigate this potential, its overhead, and to simplify the 
use of Lambda for training, testing, and evaluation machine learning models, 
we design and develop a new system and toolset called \textit{Seneca}.

Specifically, Seneca implements, packages, and deploys 
machine learning applications as stateless functions to AWS Lambda.
It then orchestrates exhaustive evaluation of specified hyperparameter settings
to identify the best performing model (for a given dataset) by
comparing the error and accuracy across models.  We consider
prediction accuracy (as opposed to explanatory power) as the
scoring metric (mean squared error for regression 
and accuracy percent for classification), to avoid overfitting.
Users present Seneca with their application, a range of values for 
each hyperparameter (or the default can be used), and a representative dataset.
Seneca produces, tests, and evaluates models for all combinations 
of hyperparameters and returns to the user
the set of parameters (or the model itself) that produces
the best cross-validation score (e.g. lowest squared error or highest
accuracy) for the dataset. Users can employ this model for other datasets
(with Seneca if desired) without retraining the model to amortize the cost
of Seneca further.

### 2019-02-03 related_work.tex
mz - spell check

The serverless framework~\cite{ref:serverless_framework} contributes an 
automated packaging and deployment toolset for serverless functions across clouds. 
The serverless framework uses Cloudformation~\cite{ref:cloudformation} to package and deploy Lambda function, which introduces additional cost. Seneca uses local docker container to avoid this cost
and overhead. The cloud infrastructure provisioning framework Terraform~\cite{ref:terraform} 
also provides automated deployment of functions to serverless platforms, but it does not simulate the AWS Lamdba runtime to create the deployment package (doing so on a non-Linux system will result in Lambda errors and failures at execution time). 
In addition, AWS CodeDeploy~\cite{ref:codedeploy} serves as component in the continuous integration and delivery pipeline for serverless architectures, but it also introduces monetary cost overhead
to do so.

>>>>

The serverless framework~\cite{ref:serverless_framework} contributes an 
automated packaging and deployment toolset for serverless functions across clouds. 
The serverless framework uses Cloudformation~\cite{ref:cloudformation} to package and deploy Lambda function, which introduces additional cost. Seneca uses local docker container to avoid this cost
and overhead. The cloud infrastructure provisioning framework Terraform~\cite{ref:terraform} 
also provides automated deployment of functions to serverless platforms, but it does not simulate the AWS Lambda runtime to create the deployment package (doing so on a non-Linux system will result in Lambda errors and failures at execution time). 
In addition, AWS CodeDeploy~\cite{ref:codedeploy} serves as component in the continuous integration and delivery pipeline for serverless architectures, but it also introduces monetary cost overhead
to do so.

### 2019-02-03 results.tex
mz - spell check

XGBoost~\cite{ref:xgboost-web} is an open source framework for gradient boosting, which 
performs both regression and classification. We consider it a regression application and
a classification application in this evaluation.  
The hyperparmeters and their defaults that Seneca uses in this study are listed in Table~\ref{tab:xgboost_para}. A definition for each hyperparameter can be found at~\cite{ref:xgboostparams}.
%xgboostparams: https://xgboost.readthedocs.io/en/latest/python/python_api.html#module-xgboost.sklearn
Support vector classification (SVC)~\cite{ref:svc} is a classification algorithm 
based on support vector machine and implemented by libsvm~\cite{ref:libsvm} library.
The hyperparmeters and their defaults that Seneca uses for SVC in this study are listed in Table~\ref{tab:svc_para}. A definition for each hyperparameter can be found at~\cite{ref:svcparams}.
%xvcparams: https://scikit-learn.org/stable/modules/generated/sklearn.svm.SVC.html
Neural network (NN)~\cite{ref:neural_network} is a machine learning framework that learns 
and make decisions from patterns in an input dataset. For this application, 
we use a feed-forward multi-layer perceptron model~\cite{ref:feedforward_nn} 
for the classification task.
The hyperparmeters and their defaults that Seneca uses for Neural Network in this study are listed in Table~\ref{tab:nn_para}. A definition for each hyperparameter can be found at~\cite{ref:nnparams}.
%nn: https://scikit-learn.org/stable/modules/generated/sklearn.neural_network.MLPClassifier.html

@@

%%%%%%%%%%%%%%%%%%%%% classification apps - tuning performance
%%%%%%%%%%%%%%%%%%%%%%%%%%
We next empirically evaluate the tuning performance of Seneca for the three
classification applications: XGBoost (classification), SVC, and NN.
Table~\ref{tab:accuracy} presents the accuracy percentage (higher is
better) reported by each application (3 right-most columns) for the default,
worst, and best hyperparameter tuning configurations reported by Seneca.
using an 80/20 (train/test) percent split,
Seneca increases accuracy by 3.43\%, 62.52\%, and 0.53\%, for XGBoost
(classification), SVC, and NN applications, respectively.  Compared to
the worst case, Seneca improves accuracy by 99.13\%, 75.75\%, and 65.57\%,
respectively.

>>>>

XGBoost~\cite{ref:xgboost-web} is an open source framework for gradient boosting, which 
performs both regression and classification. We consider it a regression application and
a classification application in this evaluation.  
The hyperparameters and their defaults that Seneca uses in this study are listed in Table~\ref{tab:xgboost_para}. A definition for each hyperparameter can be found at~\cite{ref:xgboostparams}.
%xgboostparams: https://xgboost.readthedocs.io/en/latest/python/python_api.html#module-xgboost.sklearn
Support vector classification (SVC)~\cite{ref:svc} is a classification algorithm 
based on support vector machine and implemented by libsvm~\cite{ref:libsvm} library.
The hyperparameters and their defaults that Seneca uses for SVC in this study are listed in Table~\ref{tab:svc_para}. A definition for each hyperparameter can be found at~\cite{ref:svcparams}.
%xvcparams: https://scikit-learn.org/stable/modules/generated/sklearn.svm.SVC.html
Neural network (NN)~\cite{ref:neural_network} is a machine learning framework that learns 
and make decisions from patterns in an input dataset. For this application, 
we use a feed-forward multi-layer perceptron model~\cite{ref:feedforward_nn} 
for the classification task.
The hyperparameters and their defaults that Seneca uses for Neural Network in this study are listed in Table~\ref{tab:nn_para}. A definition for each hyperparameter can be found at~\cite{ref:nnparams}.
%nn: https://scikit-learn.org/stable/modules/generated/sklearn.neural_network.MLPClassifier.html

@@

%%%%%%%%%%%%%%%%%%%%% classification apps - tuning performance
%%%%%%%%%%%%%%%%%%%%%%%%%%
We next empirically evaluate the tuning performance of Seneca for the three
classification applications: XGBoost (classification), SVC, and NN.
Table~\ref{tab:accuracy} presents the accuracy percentage (higher is
better) reported by each application (3 rightmost columns) for the default,
worst, and best hyperparameter tuning configurations reported by Seneca.
using an 80/20 (train/test) percent split,
Seneca increases accuracy by 3.43\%, 62.52\%, and 0.53\%, for XGBoost
(classification), SVC, and NN applications, respectively.  Compared to
the worst case, Seneca improves accuracy by 99.13\%, 75.75\%, and 65.57\%,
respectively.
